---
title: "Untitled"
author: "Jiaming Hu"
date: "5/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Zig-Zag (for N(10,25))

Initialization
```{r}
niter <- 10000
t.list <- NULL
theta.list <- NULL
xee.list <- NULL
gamma <- 1
# Here we consider a case where mu=10 and sigma=5, but one could generalize
# this to any N(mu,sigma^2)
mu <- 10
sigma <- 3
```

The Zig-Zag algorithm
```{r}
t.list[1] <- 0
theta.list[1] <- 1
xee.list[1] <- 1

for(k in 2:niter){
  m <- function(x){
    return(max(0,(theta.list[k-1]*(xee.list[k-1]-mu))/sigma^2 + x/sigma^2) + gamma)
  }
  
  M <- function(x,a,b){
    return(a + b*x)
  }
  
  a <- abs(theta.list[k-1]*(xee.list[k-1]-mu))/sigma^2 + gamma
  b <- 1/sigma^2
  u <- runif(1)
  s <- -log(u)
  tau <- (sqrt(a^2+2*b*s)-a)/b
  mt <- m(t.list[k-1])
  Mt <- M(t.list[k-1],a,b)
  
  t.list[k] <- t.list[k-1]+tau
  xee.list[k] <- xee.list[k-1]+theta.list[k-1]*tau
  
  if(runif(1) < m(tau)/M(tau,a,b)){
    theta.list[k] <- -theta.list[k-1]
  }else{
    theta.list[k] <- theta.list[k-1]
  }
  
}
```

```{r}
hist(xee.list,probability = T,breaks = 10,main="Normal(10,9)")
```


# Zig-Zag (for Bivariate Normal with mu = [5,10], sigma = 5I)

Initialization
```{r}
niter <- 10000
d <- 2
t.list <- NULL
theta.list <- NULL
xee.list <- NULL
gamma <- c(1,1)
mu <- c(5,10)
sigma <- 3*diag(2)
#sigma <- matrix(c(3,1,1,3),nrow=2,ncol=2)
```

The Zig-Zag algorithm
```{r}
set.seed(0521)
t.list[1] <- 0
theta.list <- rbind(theta.list,c(1,1))
xee.list <- rbind(xee.list,c(1,1))

for(k in 2:niter){
  m <- function(x,mu,sigma,theta,xee,gamma){
    m.temp <- theta*sum(sigma*(xee-mu+theta*x)) + gamma
    return(max(0,m.temp))
  }
  
  M <- function(x,a,b){
    return(a+b*x)
  }
  
  m.list <- NULL
  M.list <- NULL
  a <- NULL
  b <- NULL
  tau <- NULL
  
  for(i in 1:d){
    u <- runif(1)
    s <- -log(u)
    m.list[i] <- m(t.list[k-1],mu,sigma[i,],theta.list[k-1,i],xee.list[k-1,i],gamma[i])
    a[i] <- sum(abs(sigma[i,]*(xee.list[k-1,]-mu)) + gamma)
    b[i] <- sum(abs(sigma[i,]))
    tau[i] <- (sqrt(a[i]^2+2*b[i]*s)-a[i])/b[i]
  }
  
  i0 <- which.min(tau)
  tau <- min(tau)
  
  t.list[k] <- t.list[k-1]+tau
  xee.list <- rbind(xee.list, xee.list[k-1,]+theta.list[k-1,]*tau)
  
  mi0tau <- m(tau,mu,sigma[i0,],theta.list[k-1,i0],xee.list[k-1,i0],gamma[i0])
  Mi0tau <- M(tau,a[i0],b[i0])
  
  theta.temp <- theta.list[k-1,]
  if(runif(1) < mi0tau/Mi0tau){
    theta.temp[i0] <- -theta.temp[i0]
  }
  
  theta.list <- rbind(theta.list,theta.temp)
}
```



This is the raw points that algorithm is doing. We can clearly see the trace the
algorithm is jumping. In fact, we can see that the algoithm jumps into targeted
region fairly quickly, in about 50 iterations, with the fact that we started
from $\mu_0 = (1,1)$. This shows the Zig Zag sampler is efficient.
```{r}
aviod.0 <- plot(xee.list[,1],xee.list[,2],xlim=c(1,8),ylim=c(0,13),main="Zig Zag Sampler on Bi-variate Normal")+
  points(x=mu[1],y=mu[2],col=2,pch=19,cex=1)
```

If we use burn in, then the sample is appearing very in a very clean way.
```{r}
burn.in <- niter/10
aviod.0 <- plot(xee.list[burn.in:niter,1],xee.list[burn.in:niter,2],xlim=c(0,15),ylim=c(0,15))+
  points(x=mu[1],y=mu[2],col=2)
```



















